# Reproducible Research: Peer Assessment 1


## Loading and preprocessing the data
```{r}
setwd("~/Coursera/specdata")
data<-read.csv("activity.csv")
```

## Exclude rows where the number of steps records are na or missing
```{r}
sanitizedData<-data[complete.cases(data),]
```

# Compute the Total Number of Steps Per Day
```{r}
aggregatedSum<-aggregate(sanitizedData$steps,by=list(sanitizedData$date),sum)
```

# Histogram of  Total Number of Steps per day?
```{r}
h<-hist(aggregatedSum$x, breaks=10, col="red", xlab="Steps taken", main="Total Number of Steps Taken Per day")

```

# What is mean total number of steps taken per day?
```{r}
mean(aggregatedSum$x)
```
# What is median of total number of steps taken per day?
```{r}
median(aggregatedSum$x)
```
# Compute the Average Number of Steps Per Day By Date
```{r}
averagedDataByDate<-aggregate(sanitizedData$steps,by=list(sanitizedData$date),mean)
averagedDataByDate
```

# Compute the Average Number of Steps Per Day By Interval
```{r}
averagedDataByInterval<-aggregate(sanitizedData$steps,by=list(sanitizedData$interval),mean)
averagedDataByInterval
```



## What is the average daily activity pattern?
```{r}
 plot(averagedDataByInterval$Group.1,averagedDataByInterval$x,type="l",xlab="5 Intervals",ylab="Average Number of Steps taken")

```

## 5 Minute Interval on Average across all days which contains maximum number of steps
```{r}
IntervalWithMaxSteps<-averagedDataByInterval[averagedDataByInterval$x == max(averagedDataByInterval$x), ]

IntervalWithMaxSteps$Group.1
```


## Imputing missing values

## Total Number of Rows With Missing values In the data Set?
```{r}
 nrow(data) - nrow(sanitizedData)

```

# Replace all the missing values in the data set by the mean of steps taken per day
```{r}
 data$steps[is.na(data$steps)]<-mean(aggregatedSum$x)

```


# New Value of  Total Sum of Steps taken per day after filling in missing values 
```{r}

newaggregatedSum<-aggregate(data$steps,by=list(data$date),sum)
newaggregatedSum

```

# Updated Histogram of  Total Number of Steps per day After adding missing values?
```{r}
h2<-hist(newaggregatedSum$x, breaks=10, col="red", xlab="Steps taken", main="Total Number of Steps Taken Per day")

```

# New Value of mean of Steps taken per day after filling in missing values 
```{r}

newaggregatedmean<-aggregate(data$steps,by=list(data$date),mean)
newaggregatedmean

```

# New Value of Median of  Steps taken per day after filling in missing values 
```{r}

newaggregatedmedian<-aggregate(data$steps,by=list(data$date),median)
newaggregatedmedian

```
# New Value of Median of  Steps taken per day after filling in missing values 
Impact of imputing missing data on the data set causes the the total number of steps per date to go up higher.





## Are there differences in activity patterns between weekdays and weekends?

#Add a factor column to the data set which has the missing values filled,6 and 0 imply Sundays and Saturdays respectively, so take those as the filters
```{r}

data$datetype<-factor(as.POSIXlt(data$date)$wday)

```
#Add a levels WeekDay and Weekend to the factor cloumn added above
```{r}

levels(data$datetype)[levels(data$datetype)==0] <- "Weekend"
levels(data$datetype)[levels(data$datetype)==1] <- "Weekday"
levels(data$datetype)[levels(data$datetype)==5] <- "Weekday"
levels(data$datetype)[levels(data$datetype)==4] <- "Weekday"
levels(data$datetype)[levels(data$datetype)==3] <- "Weekday"
levels(data$datetype)[levels(data$datetype)==2] <- "Weekday"
levels(data$datetype)[levels(data$datetype)==6] <- "Weekend"
```

#The value of the new factor column datetype added based on levels specified above
stepsbyweekend

#Calculate steps taken across all weekdays
```{r}
stepsbyweekday<-subset(data,datetype=="Weekday")
stepsbyweekday
```

#Calculate steps taken across all weekends
```{r}
stepsbyweekend<-subset(data,datetype=="Weekend")
stepsbyweekend
```

## What is he average daily activity pattern?
```{r}


 plot(stepsbyweekday$interval,stepsbyweekday$steps,type="l",xlab="5 min Intervals",ylab="Avg Number of Steps taken By weekday")

 plot(stepsbyweekend$interval,stepsbyweekend$steps,type="l",xlab="5 min Intervals",ylab="Avg Number of Steps taken By weekend")

```

